### 2022-04-30 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@b6ae14646be7624385df869d6e5eb43e8f23ce1f 🚀

From the perspective of spillover risk, the most important combination
of factors is a high phylogenetic diversity of hosts with low viral
sharing; this, essentially, means that very different betacoronavirus
could co-exist within the same place. This is particularly the case
given that betacoronaviruses often evolve and even achieve host shifts
through recombination, which requires the co-occurrence of sufficiently
distinct viruses to be a major driver of emergence. In Fig. xx, this
corresponds to yellow to pale green areas, which are essentially limited
to South-Eastern Asia, and to some part of Sub-Saharan Africa. Adopting
a geographic mosaic theory perspective on risk, other regions of the
world are of lesser concern.

@@

Available data on bat betacoronavirus spillover into humans (TP overlay
on the figure) is limited and circumstantial at best for these purposes,
but our risk maps suggest that the areas predicted by prior expectations
about host biogeography correspond loosely to those where previous
emergence events have been recorded. Areas with high bat diversity and
high turnover may facilitate the evolutionary radiation of viruses,
matching previous findings that the diversification of bat coronaviruses
is driven largely by host shifts (inter-genus or higher levels of
cross-species transmission) and, to a lesser degree, cospeciation and
sharing (intra-genus cross-species transmission; Anthony et al.~2017).
This diversification - while not an actual risk factor for spillover
itself - likely increases the random chance of a virus with the raw
genomic components required for the potential to infect humans.

>>>>

From the perspective of spillover risk, the most important combination
of factors is a high phylogenetic diversity of hosts with low viral
sharing; this, essentially, means that very different betacoronavirus
could co-exist within the same place. This is particularly the case
given that betacoronaviruses often evolve and even achieve host shifts
through recombination, which requires the co-occurrence of sufficiently
distinct viruses to be a major driver of emergence. In
fig.~\ref{fig:trivariate}, this corresponds to yellow to pale green
areas, which are essentially limited to South-Eastern Asia, and to some
part of Sub-Saharan Africa. Adopting a geographic mosaic theory
perspective on risk, other regions of the world are of lesser concern
fig.~\ref{fig:risk}. Our risk decomposition does not account for viral
diversity or distinctiveness. The simple rationale behind it is that the
acquisition of viral data is rarely disconnected from the acquisition of
host data; furthermore, there are more sources of information on hosts
than on viruses, allowing to develop a host-centric perspective on risk.
Areas with high bat diversity and high turnover \emph{may} facilitate
the evolutionary radiation of viruses, matching previous findings that
the diversification of bat coronaviruses is driven largely by host
shifts (inter-genus or higher levels of cross-species transmission) and,
to a lesser degree, cospeciation and sharing (intra-genus cross-species
transmission; \textbf{Anthony et al.~2017}). This diversification is not
an actual risk factor for spillover itself, but acts downstream of a
spillover event by increasing the random chance of the emergence of a
virus with the raw genomic components required for the potential to
infect humans.

@@
